Co-citations in context

### 2019-06-04 cocit_r.tex
More revisions on 6/3/19

\emph{This paragraph needs work, but I wanted to create a placeholder for the introduction to this subsection.  It uses some of the passages form the former chacko subsection.} We have conjectured that the approach of Uzzi et al. for generating journal-pair z-scores is misspecified in its sampling from a broad dataset and its disregard for the frequency with which journal pairs are cited. Like Uzzi et al., we also used a Monte Carlo approach to simulate under a random graph model, although our principal consideration was to restrict model misspecification arising from disciplinarily irrelevant references. We addressed this consideration by analyzing disciplinary subsets of the scientific literature, thereby restricting random selection of references to only those references in the disciplinary network being studied.  In this subsection we demonstrate the effects of model misspecification or Uzzi et al.'s approach and the effectiveness of using disciplinary datasets in resolving the concomitant issues with the former.

Figure \ref{fig:scatter} shows that the z-scores for the same journal pair can be positive (negative) when computed with respect to one data set but be negative (positive) for another data set. The journal-pair z-scores in Figure \ref{fig:scatter} have consistent signs for both Immunology and WoS data sets in 71.4\% of the instances and different signs for 28.6\% of the journal pairs. When a journal-pair z-score is negative with respect to one dataset and positive with respect to another dataset, articles citing that pair are more likely to be deemed novel in the first instance and less likely to be deemed novel in the second instance.  It seems that any journal pair should either be indicative of novelty or not, but it should never have contradictory implications with a changing reference dataset. We view, therefore, these contradictions as symptomatic inappropriately sampling from a broad dataset and ignoring observed citation patterns.  Figure \ref{fig:scatter} reflects that the WoS data set has approximately 44,000 fewer negative z-scores than does the immunology data set, which contributes to its significantly lower percentage of high-novelty articles.  \emph{The plots I made of the cumulative z-score distributions for Immunology and WoS also could be used to demonstrate this phenomenon, although the scatter plot does this more clearly, and it is more striking.  The scatter plot also encodes more data, that is, the matched z-scores for each journal pair.}

Note also in Figure \ref{fig:scatter} that many z-scores change values significantly between the WoS and Immunology datasets, although the density of points near the origin in Figure \ref{fig:scatter} is significant where the z-scores may have different signs with respect to the two datasets, although their magnitudes are not significantly different. While no metric is without downside, this observation points to a weakness of measuring novelty based on the sign of the z-score where the implications of a small negative value and a large negative value are the same.

@@

Another possible symptom of model misspecification is that random sampling of citations associated with an article do not match with the distribution of citations across the disciplines observed in articles from that discipline. \citep{wallace_lariviere_gingras_2012} found that, often, a majority of an article's citations are the specialty of the article, although the percentage of citations from the same disciplinary focus varied among disciplines in the eight specialties they investigated: from approximately 39\% to 89\% in 2006. \emph{More needed here}  Similarly, we found that 77.4\% of citation in applied physics articles in 1985 (replace with 1995 data?) were from applied physics.  \emph{Add references to other datasets.} A random sampling of citations would ideally reflect the percentages from the discipline in question, as well as the percentages of articles from other discipline.

@@

\emph{Insert comment about the left-hand side of Figure \ref{fold_effect} where the disciplinary dataset has resolved the fold issue.}  We also predicted that model misspecification as measured by the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and simulated frequencies in a disciplinary network would be less than the divergence for observed and simulated frequencies in the WoS superset. The results in Table \ref{tab:KLD} indicate that for the set of journals common to both a disciplinary network and the WoS superset, simulations under our model consistently have a lower K-L divergence compared to simulations that draw from the WoS superset (and its attendant substitutions that are ectopic with respect to field and discipline).

\begin{figure}
\label{fold_effect}
\centering
\includegraphics[width=0.5\hsize]{fold_effect.pdf}
\caption{Intra-network citation shuffling preserves disciplinary composition. Publications of type Article belonging to the WoS research areas (i) Arts and Humanities (ii) Life Sciences and Biomedicine were subject to a single shuffle of all their cited references using either the cited references in these networks as a source of random substitutions (bg\_local) or references from all articles in WoS (bg\_allWoS). The disciplinary composition of cited reference before and after shuffling was measured for each of 151 sub-groups (from the extended subject classification in WoS) in these two research areas, expressed as fold difference between citation counts grouped by subject for original (o) and shuffled (s) references using the formula (fold\_difference = $ifelse(o > s, o/s, s/o)$) and rounded to the nearest integer. A fold difference of 1 indicates that citation shuffling did not alter disciplinary composition. Data are shown for articles published in 1985. All four boxplots are generated from 151 observations. Min \& Max values for fold differences were as follows. (i) Arts \& Humanities. bg\_local (1,2), bg\_allWoS (1, 488) (ii) Life Sciences \& Biomedicine. bg\_local(1,1), bg\_allWos (1, 352) Note y-axis: $log_2$ scale.} 
\end{figure}

>>>>

\emph{This paragraph needs work, but I wanted to create a placeholder for the introduction to this subsection.  It uses some of the passages form the former Chacko subsection.} We have conjectured that the approach of Uzzi et al. for generating journal-pair z-scores is misspecified in its sampling from a broad dataset and its disregard for the frequency with which journal pairs are cited. Like Uzzi et al., we also used a Monte Carlo approach to simulate under a random graph model, although our principal consideration was to restrict model misspecification arising from disciplinarily irrelevant references. We addressed this consideration by analyzing disciplinary subsets of the scientific literature, thereby restricting random selection of references to only those references in the disciplinary network being studied.  In this subsection we demonstrate the effects of model misspecification or Uzzi et al.'s approach and the effectiveness of using disciplinary datasets in resolving the concomitant issues with the former.

Figure \ref{fig:scatter} shows that the z-scores for the same journal pair can be positive (negative) when computed with respect to one data set but be negative (positive) for another data set. The journal-pair z-scores in Figure \ref{fig:scatter} have consistent signs for both Immunology and WoS data sets in 71.4\% of the instances and different signs for 28.6\% of the journal pairs. When a journal-pair z-score is negative with respect to one dataset and positive with respect to another dataset, articles citing that pair are more likely to be deemed novel in the first instance and less likely to be deemed novel in the second instance.  It seems that any journal pair should either be indicative of novelty or not, but it should never have contradictory implications that depend on the reference dataset. We view, therefore, these contradictions as symptomatic of inappropriately sampling from a broad dataset and ignoring observed citation patterns.  Figure \ref{fig:scatter} reflects that the WoS data set has approximately 44,000 fewer negative z-scores than does the immunology data set, which contributes to its significantly lower percentage of high-novelty articles.  \emph{The plots I made of the cumulative z-score distributions for Immunology and WoS also could be used to demonstrate this phenomenon, although the scatter plot does this more clearly, and it is more striking.  The scatter plot also encodes more data, that is, the matched z-scores for each journal pair.}

Note also in Figure \ref{fig:scatter} that many z-scores values are significantly different between the WoS and Immunology datasets, although the density of points near the origin in Figure \ref{fig:scatter} indicates that some journal pairs change signs between the datasets while their magnitudes are not significantly different. While no metric is without downside, this observation points to a weakness of measuring novelty based simply on the sign of z-scores where no distinction is made between a small negative value and a large negative value, but a negative value of small magnitude is viewed as being significantly different than a small positive value.

@@

Another symptom of model misspecification is that random sampling of citations associated with an article do not match with the distribution of citations across the disciplines observed in articles from that discipline. \citep{wallace_lariviere_gingras_2012} found that, often, a majority of an article's citations are from the specialty of the article, although that percentage varied among disciplines in the eight specialties they investigated: from approximately 39\% to 89\% for 2006. \emph{More needed here.}  Similarly, we found that 77.4\% of citations in applied physics articles in 1985 (replace with 1995 data?) were from applied physics.  \emph{Add references to other datasets.} A random sampling of citations would ideally reflect the percentages from the discipline in question, as well as the percentages of articles from other disciplines.

@@

\emph{Insert comment about the left-hand side of Figure \ref{fold_effect} where the disciplinary dataset has resolved the fold issue.}  We also predicted that model misspecification as measured by the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and simulated frequencies in a disciplinary network would be less than the divergence for observed and simulated frequencies in the WoS superset. The results in Table \ref{tab:kld} indicate that for the set of journals common to both a disciplinary network and the WoS superset, simulations under our model consistently have a lower K-L divergence compared to simulations that draw from the WoS superset (and its attendant substitutions that are ectopic with respect to field and discipline).

\begin{figure}
\centering
\includegraphics[width=0.5\hsize]{fold_effect.pdf}
\caption{\label{fold_effect} Intra-network citation shuffling preserves disciplinary composition. Publications of type Article belonging to the WoS research areas (i) Arts and Humanities (ii) Life Sciences and Biomedicine were subject to a single shuffle of all their cited references using either the cited references in these networks as a source of random substitutions (bg\_local) or references from all articles in WoS (bg\_allWoS). The disciplinary composition of cited reference before and after shuffling was measured for each of 151 sub-groups (from the extended subject classification in WoS) in these two research areas, expressed as fold difference between citation counts grouped by subject for original (o) and shuffled (s) references using the formula (fold\_difference = $ifelse(o > s, o/s, s/o)$) and rounded to the nearest integer. A fold difference of 1 indicates that citation shuffling did not alter disciplinary composition. Data are shown for articles published in 1985. All four boxplots are generated from 151 observations. Min \& Max values for fold differences were as follows. (i) Arts \& Humanities. bg\_local (1,2), bg\_allWoS (1, 488) (ii) Life Sciences \& Biomedicine. bg\_local(1,1), bg\_allWos (1, 352) Note y-axis: $log_2$ scale.} 
\end{figure}
